---
title       : rCharts using morris.js
subtitle    : 
author      : TimelyPortfolio
framework   : minimal       # {io2012, html5slides, shower, dzslides, ...}
github: {user: timelyportfolio, repo: rCharts_morris_standalone, branch: "gh-pages"}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : [morris]      # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
---

## rCharts morris.js Examples
morris.js offers some of the best time series handling of all the js libraries provided through rCharts.  This makes it ideal for much of the time-series financial analysis performed in **PerformanceAnalytics**.

### Cumulative Growth Line Chart
As a first example, let's build a cumulative growth line chart.  It takes only a few lines of code to produce the chart.
```{r,warning=FALSE,message=FALSE,results="asis",tidy.opts=list(width.cutoff=50)}
#initial setup
require(rCharts)
require(reshape2)
require(PerformanceAnalytics)

data(edhec)

#nice with Morris that we can do with melt or without
edhec.df <- data.frame(format( index(edhec), "%Y-%m-%d" ),
                       apply( 1 + coredata(edhec), MARGIN = 2, FUN = cumprod ),
                       stringsAsFactors = FALSE)
colnames(edhec.df) <- c("date",
                        gsub(x=colnames(edhec),
                             pattern=" ",
                             replacement="") )

m1a <- mPlot( x = "date",
              y = colnames(edhec.df)[2:NCOL(edhec.df)],
              data = edhec.df,
              type = "Line")
m1a$set ( pointSize = 0)
m1a$set ( hideHover = "auto" )
m1a$print()
```

### Grouped Bar of Annual Returns by Year
As another example, let's look at a grouped bar chart plotting the annual return for each year by manager/index.
```{r,warning=FALSE,message=FALSE,results="asis",tidy.opts=list(width.cutoff=50)}
data(managers)

#get calendar year returns so starting at 13
#monthly returns are 1 through 12
perf <- data.frame(table.CalendarReturns(managers[,c(1,8,9)])[,13:(12 + NCOL(managers[,c(1,8,9)]))],
                   stringsAsFactors = FALSE)

#make rownames a column
perf <- data.frame(rownames( perf ), perf, row.names = NULL)
#add date to column names and remove . from column names to not confuse js
colnames( perf ) <- c("date",
                      gsub(x = colnames( perf[-1] ),
                         pattern = "[.]",
                         replacement = "" ) )

#build the plot
m2 <- mPlot( x = "date", 
             y = colnames( perf )[-1],
             data = perf,
             type = "Bar")
#not pretty colors but an example how we can specify
m2$set(barColors = brewer.pal(10, "BrBG")[c(8,4,5)])
m2$set(postUnits = "%")
m2$set(hideHover = "auto")
m2$print()
```


### Grouped Bar Return Statistics
Another way to use a grouped bar in a manager performance setting might be a look at cumulative or rolling performance statistics.  In this case, the x-axis will be an ordinal scale instead of a date-time scale.
```{r,warning=FALSE,message=FALSE,results="asis",tidy.opts=list(width.cutoff=50)}
data(managers)

#get columns 3 through 12 and 14 as statistics with similar ranges
pastats <- data.frame(table.Stats(managers[,c(1,8,9)])[c(3:12,14),],
                   stringsAsFactors = FALSE) * 100

#make rownames a column
pastats <- data.frame(rownames( pastats ), pastats, row.names = NULL)
#add date to column names and remove . from column names to not confuse js
colnames( pastats ) <- c("metric",
                      gsub(x = colnames( pastats[-1] ),
                           pattern = "[.]",
                           replacement = "" ) )

#build the plot
m3 <- mPlot( x = "metric", 
             y = colnames( pastats )[-1],
             data = pastats,
             type = "Bar",
             width = 1000)

#not pretty colors but an example how we can specify
m3$set(barColors = brewer.pal(9, "PuBuGn")[c(7,6,3)])
#set to add % at end
m3$set(postUnits = "%")
#set to always show hover
m3$set(hideHover = 'auto')
m3$print()
```

### Comparison to PerformanceAnalytics Charts
These were minimal examples using no additional javascript programming.  rCharts allows this capability, and I strongly encourage readers to check out these two examples from the package author Ramnath Vaidyanathan [NYT Baseball Strikeout Tutorial](http://ramnathv.github.io/rChartsNYT/) and [Visualizing the Reinhart and Rogoff Public Debt Study](http://glimmer.rstudio.com/ramnathv/rChartsRogoff/).  I will build on these examples to add additional interactivity.

In a future session I will compare the javascript interactive plots to the static graphics from R.  I think most will agree that for most cases an interactive presentation will allow much better analysis and communication.